---
title: "NCEAS_test"
output:  
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(readr)
```

# Introduction

I downloaded the file [BGchem2008data.csv](https://arcticdata.io/catalog/view/doi:10.18739/A25T3FZ8X) from the Arctic Data Center (doi:10.18739/A25T3FZ8X)

1. read in data

```{r chunkerooni, message=FALSE}
bg_chem <- read_csv("BGchem2008data.csv")
  
```

# Analysis

We are interested in nutrient concentrations over the whole water column, so we calculated the mean value of nutrients of interest (nitrate, nitrite, ammonium, and phosphate)

```{r sumstats}

nitrate <- mean(bg_chem$NO3)
nitrite <- mean(bg_chem$NO2)
amm <- mean(bg_chem$NH4)
phos <- mean(bg_chem$P)

```

1. calculate mean Redfield ratio

```{r}

ratio <- (nitrate + nitrite + amm) / phos
```

We calculated the Redfield ratio to be `r round(ratio)`

1. plot Redfield ratio

```{r, echo = FALSE}

plot(bg_chem$P, bg_chem$NO2 + bg_chem$NO3 + bg_chem$NH4)

```

