---
title: "Session 8"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, message = FALSE}

library(dplyr)
library(tidyr)
library(readr)

```

## Data Sources

### read in the data

Use read_csv from readr package because it can read data directly from the web.  Apparently there is a google drive package to load things in directly from google drive.  These are ways to make the R markdown self-contained.

```{r load data}

catch_original <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")

```

## Clean and Reshape data

### remove unnecessary columns

```{r}

catch_data <- catch_original %>% 
  select(-All, -notesRegCode)

```
### check column typing

```{r}

catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook== "I", 1, Chinook)) %>%
  mutate(Chinook = as.numeric(Chinook))

```

### reshape data

gather was renamed to pivot_longer and pivot_wider in tidyr

```{r}

catch_long <- catch_clean %>% 
  pivot_longer(-c(Region, Year), names_to = "species", values_to = "catch")

catch_wide <- catch_long %>% 
  pivot_wider(names_from = "species", values_from = "catch")

#rename unit of catch
catch_long <- catch_long %>% 
  rename(catch_thousands = catch) 

catch_long <- catch_long %>% 
  mutate(catch = catch_thousands*1000) %>% 
  select(-catch_thousands)

```

```{r example summary stats}

mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(catch_mean = mean(catch)) %>% 
  arrange(catch_mean)

n_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(n = n())

SSE_catch <- catch_long %>% 
  filter(Region == "SSE")

```

## Join to Regions dataset

```{r load regions data, message=FALSE}

region_defs <- read_csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
  select(code, mgmtArea)

```

